---
title: "Mixture model for the sample"
author: "Marc Comas-Cufí,  Antonella Buccianti, Josep Antoni Martín-Fernández and Glòria Mateu-Figueras"
date: "July 16, 2015"
output: 
  html_document: 
    theme: readable
    toc: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, echo=FALSE, comment="   ")
options(width=200)
library(ggmap)
library(ggplot2)
library(tidyr)
library(dplyr)
library(mixpack)
library(gridExtra)
library(mclust)
load('data/iamg_data.RData')
load('data/iamg_map.RData')
```

```{r}
clean_zeros = function(.data) .data  %>% subset(apply(.data %>% select(HCO3, Ca, Cl, Mg, K, Na, SO4), 1, prod) != 0)
clean_non_located = function(.data) .data  %>% subset(complete.cases(.data %>% select(est, nord)))
compo = function(.data) .data %>% select(HCO3, Ca, Cl, Mg, K, Na, SO4) %>% subset(apply(., 1, prod) != 0)
mgL_to_mmol = function(.data) .data %>% mutate(
  HCO3=HCO3/61.02,
  Ca=Ca/40.078,
  Cl=Cl/35.4527,
  Mg=Mg/24.305,
  K=K/39.0983,
  Na=Na/22.989,
  SO4=SO4/96.06 )
mmol_to_meq = function(.data) .data %>% mutate(
  HCO3=HCO3,
  Ca=2*Ca,
  Cl=Cl,
  Mg=2*Mg,
  K=K,
  Na=Na,
  SO4=2*SO4
)
tern_ca = function(.data) .data %>% 
  mutate(Ca = Ca, Mg = Mg, Na.K = Na + K) %>% 
  select(Ca, Mg, Na.K)
tern_an = function(.data) .data %>% 
  select(Cl, SO4, HCO3)
closure = function(.data, k = 100) (k*.data/apply(.data, 1, sum)) %>% round(4) %>% tbl_df
gmean = function(...){
  .data = data.frame(list(...))
  apply(.data, 1, prod)^(1/NCOL(.data))
}
balance = function(.data) .data %>% select(HCO3, Ca, Cl, Mg, K, Na, SO4) %>%
  mutate(
    b1 = sqrt(4*3/(4+3)) * log(gmean(Ca, Mg, Na, K) / gmean(HCO3, SO4, Cl)),
    b2 = sqrt(2*2/(2+2)) * log(gmean(Ca, Mg) / gmean(Na, K)),
    b3 = sqrt(1*1/(1+1)) * log(Ca/Mg),
    b4 = sqrt(1*1/(1+1)) * log(Na/K),
    b5 = sqrt(1*2/(1+2)) * log(gmean(HCO3)/gmean(SO4, Cl)),
    b6 = sqrt(1*1/(1+1)) * log(SO4/Cl) ) %>% select(b1,b2,b3,b4,b5,b6)
```


```{r}
data = data %>% clean_zeros %>% clean_non_located
```

If we try to sampling from the original dataset. It is likely to obtain sample from location where the where a higher amount of extraction. To avoid sampling always from the same regions we applied an [inverse probability bootstrap](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC4488419/) strategy by first, grouping the data using a hierarchical parition (HC) algorithm and then calculating the probability of being chosen by calculating the size of the HC cluster where data lied.

```{r, fig.height=3, fig.width=3, results='hold'}
N = 1000
ALPHA=0.05
HCLEVEL = 300
set.seed(1)
id_sample = sample(1:nrow(data), N)

data.sim = data %>% slice(id_sample)
ggmap(map) + 
  geom_point(data = data.sim, aes(x = lon, y = lat), alpha=ALPHA) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank()) + ggtitle('Direct sampling')

## Hierarchical partition
res = hclust(data %>% select(est, nord) %>% dist, method="ward.D2")
cluster = cutree(res, HCLEVEL)
(p1<-ggmap(map) + 
  geom_point(data = data %>% mutate(g=as.factor(cluster)), 
             aes(x = lon, y = lat, col=g), alpha=1) +
  xlab(NULL) + ylab(NULL) +
  theme(legend.position = 'none', 
        axis.ticks = element_blank(), 
        axis.text = element_blank())) + ggtitle('Hierarchical partition')
#p1 + coord_cartesian(xlim= c(10.75, 11.25), ylim = c(43.75, 44.25))

# IPW
data.p = data %>% mutate(
  g = cluster,
  n.g = table(g)[g],
  p.g = n.g/nrow(data),
  w = 1/p.g,
  p = w/sum(w))
data.sim = data.p %>% slice(sample(1:nrow(data.p), N, prob = data.p$p, replace = FALSE))
ggmap(map) + 
  geom_point(data = data.sim, aes(x = lon, y = lat), alpha=ALPHA) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank()) + ggtitle('IPW sampling')
#p2 + coord_cartesian(xlim= c(10.75, 11.25), ylim = c(43.75, 44.25))
```

Final sample:

```{r}
(d.meq <- data.sim %>% compo %>% mgL_to_mmol %>% mmol_to_meq)
```

## CLR-biplot

```{r, fig.width=9, fig.height=7}
source('R/plot_functions.R')
alpha = .25
Xclr = clr_coordinates(X = d.meq) %>% tbl_df
labels = sprintf("clr(%s)", names(Xclr))

grid.arrange(
  ggplot_biplot(X = Xclr, labels = labels, x = 1, y = 2, alpha=alpha),
  ggplot_biplot(X = Xclr, labels = labels, x = 3, y = 2, alpha=alpha),
  ggplot_biplot(X = Xclr, labels = labels, x = 1, y = 3, alpha=alpha), 
  nrow = 2, ncol = 2, main='CLR-biplot')
```

# Mixture models using balances

We are going to work with following balances:

```
b1 = sqrt(4*3/(4+3)) * log(gmean(Ca, Mg, Na, K) / gmean(HCO3, SO4, Cl)),
b2 = sqrt(2*2/(2+2)) * log(gmean(Ca, Mg) / gmean(Na, K)),
b3 = sqrt(1*1/(1+1)) * log(Ca/Mg),
b4 = sqrt(1*1/(1+1)) * log(Na/K),
b5 = sqrt(1*2/(1+2)) * log(gmean(HCO3)/gmean(SO4, Cl)),
b6 = sqrt(1*1/(1+1)) * log(SO4/Cl) 
```
## Working with the full composition

```{r, fig.width=9, fig.height=9}
K = 9
mod <- Mclust(d.meq %>% balance, K)
plot.Mclust(mod, col = rainbow(K), what = 'classification')
```

```{r, fig.width=9, fig.height=9}
#data.post = cbind(data.sim %>%mutate(clust=as.factor(mod$classification)), 
#                  as.data.frame(mod$z))  %>% gather(key=comp, value=post, V1:V9)

data.post = cbind(data.sim %>% mutate(clust=as.factor(mod$classification)), 
                  data.frame(post = apply(mod$z, 1, max)))  %>% tbl_df

ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post, col=clust), size=4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~clust, nrow=3) +
  scale_colour_manual(values=rainbow(K)) + 
  ggtitle('IPW sampling')
```

## Separating cations and anions

 * Cations
 
```{r, fig.width=3, fig.height=3, results='hold'}
mod.cat <- Mclust(d.meq %>% balance %>% select(b2, b3, b4),3)
lbls = c('Ca', 'Mg', 'Na', 'K')
ggplot_biplot(clr_coordinates( d.meq %>% select(one_of(lbls))), 
              labels = sprintf('clr_%s', lbls), x = 1, y = 2, 
              col = rainbow(3)[mod.cat$classification], alpha=1) + theme(legend.position='none')
ggplot_biplot(clr_coordinates( d.meq %>% select(one_of(lbls))), 
              labels = sprintf('clr_%s', lbls), x = 1, y = 3, 
              col = rainbow(3)[mod.cat$classification], alpha=1) + theme(legend.position='none')
```

 * Anions
 
```{r, fig.width=3, fig.height=3}
mod.an <- Mclust(d.meq %>% balance %>% select(b5, b6), 5)
lbls = c('Cl', 'SO4', 'HCO3')
ggplot_biplot(clr_coordinates( d.meq %>% select(one_of(lbls))), 
              labels = sprintf('clr_%s', lbls), x = 1, y = 2, 
              col = rainbow(5)[mod.an$classification], alpha=1) + theme(legend.position='none')
```

```{r, fig.width=9, fig.height=7}
#data.post = cbind(data.sim %>%mutate(clust=as.factor(mod$classification)), 
#                  as.data.frame(mod$z))  %>% gather(key=comp, value=post, V1:V9)
data.post = cbind(data.sim %>% mutate(clust.cat=as.factor(mod.cat$classification)), 
                  data.frame(post.cat = apply(mod.cat$z, 1, max),
                             clust.an = as.factor(mod.an$classification),
                             post.an=  apply(mod.an$z, 1, max))) %>% mutate(
                               facie = sprintf('%d-%d', clust.cat, clust.an)
                             ) %>% tbl_df
ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post.cat*post.an, col=facie), size=4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_grid(clust.cat~clust.an) +
  scale_colour_manual(values=rainbow(15)) + 
  ggtitle('IPW sampling')
```

## A hierarchical clustering 

```{r}
hp = get_hierarchical_partition(tau = mod$z,
                                omega = function(v_tau, a) v_tau[a],
                                lambda = function(v_tau, a, b) log(v_tau[b]/v_tau[a])^2)
ggmap(map) + 
  geom_point(data = data.post %>% mutate(g = cluster_partition(mod$z, hp[[6]])), 
             aes(x = lon, y = lat), size=4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~g) +
  scale_colour_manual(values=rainbow(15)) + 
  ggtitle('IPW sampling')
```