---
title: "Mixture model for the sample using the full composition"
author: "Marc Comas-Cufí,  Antonella Buccianti, Josep Antoni Martín-Fernández and Glòria Mateu-Figueras"
date: "July 16, 2015"
output: 
  html_document: 
    theme: readable
    toc: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, echo=FALSE, comment="   ")
options(width=220)
library(ggmap)
library(ggplot2)
library(tidyr)
library(dplyr)
library(mixpack)
library(gridExtra)
library(mclust)
source('R/plot_functions.R')
load('data/iamg_data.RData')
load('data/iamg_map.RData')
```

```{r}
clean_zeros = function(.data) .data  %>% subset(apply(.data %>% select(HCO3, Ca, Cl, Mg, K, Na, SO4), 1, prod) != 0)
clean_non_located = function(.data) .data  %>% subset(complete.cases(.data %>% select(est, nord)))
compo = function(.data) .data %>% select(HCO3, Ca, Cl, Mg, K, Na, SO4) %>% subset(apply(., 1, prod) != 0)
mgL_to_mmol = function(.data) .data %>% mutate(
  HCO3=HCO3/61.02,
  Ca=Ca/40.078,
  Cl=Cl/35.4527,
  Mg=Mg/24.305,
  K=K/39.0983,
  Na=Na/22.989,
  SO4=SO4/96.06 )
mmol_to_meq = function(.data) .data %>% mutate(
  HCO3=HCO3,
  Ca=2*Ca,
  Cl=Cl,
  Mg=2*Mg,
  K=K,
  Na=Na,
  SO4=2*SO4
)
tern_ca = function(.data) .data %>% 
  mutate(Ca = Ca, Mg = Mg, Na.K = Na + K) %>% 
  select(Ca, Mg, Na.K)
tern_an = function(.data) .data %>% 
  select(Cl, SO4, HCO3)
closure = function(.data, k = 100) (k*.data/apply(.data, 1, sum)) %>% round(4) %>% tbl_df
gmean = function(...){
  .data = data.frame(list(...))
  apply(.data, 1, prod)^(1/NCOL(.data))
}
balance = function(.data) .data %>% select(HCO3, Ca, Cl, Mg, K, Na, SO4) %>%
  mutate(
    b1 = sqrt(4*3/(4+3)) * log(gmean(Ca, Mg, Na, K) / gmean(HCO3, SO4, Cl)),
    b2 = sqrt(2*2/(2+2)) * log(gmean(Ca, Mg) / gmean(Na, K)),
    b3 = sqrt(1*1/(1+1)) * log(Ca/Mg),
    b4 = sqrt(1*1/(1+1)) * log(Na/K),
    b5 = sqrt(1*2/(1+2)) * log(gmean(HCO3)/gmean(SO4, Cl)),
    b6 = sqrt(1*1/(1+1)) * log(SO4/Cl) ) %>% select(b1,b2,b3,b4,b5,b6)
balance_nms = c('b1' = '(Ca·Mg·Na·K)/(HCO3·SO4·Cl)',
                'b2' = '(Ca·Mg)/(Na·K)',
                'b3' = 'Ca/Mg',
                'b4' = 'Na/K',
                'b5' = 'HCO3/(SO4·Cl)', 
                'b6' = 'SO4·Cl')

base = list(c(1, 1, 1, 1,-1,-1,-1),
            c(1, 1,-1,-1, 0, 0, 0),
            c(1,-1, 0, 0, 0, 0, 0),
            c(0, 0, 1,-1, 0, 0, 0),
            c(0, 0, 0, 0, 1,-1,-1),
            c(0, 0, 0, 0, 0, 1,-1))
pos = function(n,d) 1/n * sqrt(n*d/(n+d))
neg = function(n,d) -1/d * sqrt(n*d/(n+d))
base = lapply(base, function(x){
  res = x
  res[x == 1] = pos(sum(x==1), sum(x==-1))
  res[x == -1] = neg(sum(x==1), sum(x==-1))
  res[x == 0] = 0
  exp(res) / sum(exp(res))
})
balance_coda = function(Xbal){
  X = Reduce('*', lapply(1:NCOL(Xbal), function(i) sapply(base[[i]], function(x) x^Xbal[,i])))
  (X/apply(X,1,sum)) %>% data.frame %>% setNames(c('Ca','Mg', 'Na', 'K', 'HCO3', 'SO4', 'Cl'))
}
```


```{r}
data = data %>% clean_zeros %>% clean_non_located

N = 1000
ALPHA=0.05
HCLEVEL = 300
set.seed(1)

data.sim = data %>% slice(sample(1:nrow(data), N, replace = TRUE))
```

We start with aa random sample

```{r}
(d.meq <- data.sim %>% compo %>% mgL_to_mmol %>% mmol_to_meq)
```

# Mixture models

We work with following balances

We are going to work with following balances:

```
b1 = sqrt(4*3/(4+3)) * log(gmean(Ca, Mg, Na, K) / gmean(HCO3, SO4, Cl)),
b2 = sqrt(2*2/(2+2)) * log(gmean(Ca, Mg) / gmean(Na, K)),
b3 = sqrt(1*1/(1+1)) * log(Ca/Mg),
b4 = sqrt(1*1/(1+1)) * log(Na/K),
b5 = sqrt(1*2/(1+2)) * log(gmean(HCO3)/gmean(SO4, Cl)),
b6 = sqrt(1*1/(1+1)) * log(SO4/Cl) 
```

```{r, fig.width=9, fig.height=9}
K = 15
summary(mod <- Mclust(d.meq %>% balance, K))
```

How are the mean of each component?

Mean representation for each component (using balances)

```{r, fig.width=12, fig.height=6}
df = mod$parameters$mean %>% t %>% data.frame %>% setNames(balance_nms) %>%
  mutate(id = sprintf('%02d', 1:K)) %>% 
  gather(key='balance', value='value', -id)
df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=id, y=value, colour=id), size=4) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=3, scale='free') + 
  theme_bw() + theme(legend.position = 'none')
```

Center for each component

```{r, fig.width=12, fig.height=6}
df = mod$parameters$mean %>% t %>% data.frame %>% balance_coda
df$id = sprintf('%02d', 1:K)  
df = df %>% gather(key='comp', value='value', -id)
df.mean = df %>% group_by(comp) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=comp, y=value, colour=id), size=4) +
  facet_wrap(~id, ncol=3) + 
  theme_bw() + theme(legend.position = 'none')
```

The centers in a biplot

```{r}
df = mod$parameters$mean %>% t %>% data.frame %>% balance_coda
Xclr = clr_coordinates(X = df) %>% tbl_df
labels = sprintf("clr(%s)", names(Xclr))


ggplot_biplot(X = Xclr, labels = labels, x = 1, y = 2, alpha=0.9)
```

## CLR-biplot

The sample represented in a biplot colored clustered with respect the component

```{r, fig.width=9, fig.height=7}
source('R/plot_functions.R')
alpha = .95
Xclr = clr_coordinates(X = d.meq) %>% tbl_df
labels = sprintf("clr(%s)", names(Xclr))

grid.arrange(
  ggplot_biplot(X = Xclr, labels = labels, x = 1, y = 2, alpha=alpha, 
                col = as.character(mod$classification)) + 
    theme(legend.position='none') + scale_colour_manual(values=rainbow(K)),
  ggplot_biplot(X = Xclr, labels = labels, x = 3, y = 2, alpha=alpha,
                col = as.character(mod$classification)) + 
    theme(legend.position='none') + scale_colour_manual(values=rainbow(K)),
  ggplot_biplot(X = Xclr, labels = labels, x = 1, y = 3, alpha=alpha,
                col = as.character(mod$classification)) + 
    theme(legend.position='none') + scale_colour_manual(values=rainbow(K)), 
  nrow = 2, ncol = 2, main='CLR-biplot')
```

# Hierarchical merging

```{r, include=F}
xlog = function(x) x * log(x)
l_lambda = list(
  'entr' = function(v_tau, a, b) xlog(v_tau[a] + v_tau[b]) - xlog(v_tau[a]) - xlog(v_tau[b]),
  'demp' = function(v_tau, a, b) if(which.max(v_tau) == b) 1 else 0,
  'demp.mod' = function(v_tau, a, b) v_tau[b] * (v_tau[a] + v_tau[b])^-1,
  'coda' = function(v_tau, a, b) log(v_tau[b] / v_tau[a]),
  'coda.norm' = function(v_tau, a, b) -log(v_tau[b] / v_tau[a])^2,
  'prop' = function(v_tau, a, b) v_tau[b] )

# Weitghing functions
l_omega = list(
  'cnst' = function(v_tau, a) 1,
  'prop' = function(v_tau, a) v_tau[a],
  'dich' = function(v_tau, a) if(which.max(v_tau) == a) 1 else 0
)
```

```{r}
POST = mod$z#[,-c(6,12)]
POST[POST==0] = .Machine$double.xmin
hp = get_hierarchical_partition(POST,
                                omega = l_omega[['prop']],
                                lambda = l_lambda[['coda.norm']])
hp.mod = hp
```

## 15 components

```{r, results='hold', fig.width=4.4, fig.height=8}
hp_temp = hp.mod[[15]]
data.post = cbind(data.sim %>% mutate(clust=cluster_partition(mod$z, hp_temp)), 
                  data.frame(post = apply(mod$z, 1, max)))  %>% tbl_df
print(ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post), size=4, alpha=0.2) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~clust, ncol=3) +
  scale_colour_manual(values=rainbow(K)) )

df = mod$parameters$mean %>% t %>% data.frame %>% setNames(balance_nms) %>%
  mutate(id = 1:K) %>% 
  gather(key='balance', value='value', -id)
for(i in names(hp_temp)){
  df$id[df$id %in% hp_temp[[i]]] = i
}
df$id = stringr::str_sub(df$id, 1, 5)

df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=id, y=value, colour=id), size=2) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90))
```

## 14 components

```{r, results='hold', fig.width=4.4, fig.height=8}
hp_temp = hp.mod[[14]]
data.post = cbind(data.sim %>% mutate(clust=cluster_partition(mod$z, hp_temp)), 
                  data.frame(post = apply(mod$z, 1, max)))  %>% tbl_df
print(ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post), size=4, alpha=0.2) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~clust, ncol=3) +
  scale_colour_manual(values=rainbow(K)) )

df = mod$parameters$mean %>% t %>% data.frame %>% setNames(balance_nms) %>%
  mutate(id = 1:K) %>% 
  gather(key='balance', value='value', -id)
for(i in names(hp_temp)){
  df$id[df$id %in% hp_temp[[i]]] = i
}
df$id = stringr::str_sub(df$id, 1, 5)

df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=id, y=value, colour=id), size=2) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90))
```

## 13 components

```{r, results='hold', fig.width=4.4, fig.height=8}
hp_temp = hp.mod[[13]]
data.post = cbind(data.sim %>% mutate(clust=cluster_partition(mod$z, hp_temp)), 
                  data.frame(post = apply(mod$z, 1, max)))  %>% tbl_df
print(ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post), size=4, alpha=0.2) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~clust, ncol=3) +
  scale_colour_manual(values=rainbow(K)) )

df = mod$parameters$mean %>% t %>% data.frame %>% setNames(balance_nms) %>%
  mutate(id = 1:K) %>% 
  gather(key='balance', value='value', -id)
for(i in names(hp_temp)){
  df$id[df$id %in% hp_temp[[i]]] = i
}
df$id = stringr::str_sub(df$id, 1, 5)

df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=id, y=value, colour=id), size=2) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90))
```

## 12 Components

```{r, results='hold', fig.width=4.4, fig.height=8}
hp_temp = hp.mod[[12]]
data.post = cbind(data.sim %>% mutate(clust=cluster_partition(mod$z, hp_temp)), 
                  data.frame(post = apply(mod$z, 1, max)))  %>% tbl_df
print(ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post), size=4, alpha=0.2) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~clust, ncol=3) +
  scale_colour_manual(values=rainbow(K)) )

df = mod$parameters$mean %>% t %>% data.frame %>% setNames(balance_nms) %>%
  mutate(id = 1:K) %>% 
  gather(key='balance', value='value', -id)
for(i in names(hp_temp)){
  df$id[df$id %in% hp_temp[[i]]] = i
}
df$id = stringr::str_sub(df$id, 1, 5)

df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=id, y=value, colour=id), size=2) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90))
```

## 11 components

```{r, results='hold', fig.width=4.4, fig.height=8}
hp_temp = hp.mod[[11]]
data.post = cbind(data.sim %>% mutate(clust=cluster_partition(mod$z, hp_temp)), 
                  data.frame(post = apply(mod$z, 1, max)))  %>% tbl_df
print(ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post), size=4, alpha=0.2) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~clust, ncol=3) +
  scale_colour_manual(values=rainbow(K)) )

df = mod$parameters$mean %>% t %>% data.frame %>% setNames(balance_nms) %>%
  mutate(id = 1:K) %>% 
  gather(key='balance', value='value', -id)
for(i in names(hp_temp)){
  df$id[df$id %in% hp_temp[[i]]] = i
}
df$id = stringr::str_sub(df$id, 1, 5)

df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=id, y=value, colour=id), size=2) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90))
```

## 10 components

```{r, results='hold', fig.width=4.4, fig.height=8}
hp_temp = hp.mod[[10]]
data.post = cbind(data.sim %>% mutate(clust=cluster_partition(mod$z, hp_temp)), 
                  data.frame(post = apply(mod$z, 1, max)))  %>% tbl_df
print(ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post), size=4, alpha=0.2) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~clust, ncol=3) +
  scale_colour_manual(values=rainbow(K)) )

df = mod$parameters$mean %>% t %>% data.frame %>% setNames(balance_nms) %>%
  mutate(id = 1:K) %>% 
  gather(key='balance', value='value', -id)
for(i in names(hp_temp)){
  df$id[df$id %in% hp_temp[[i]]] = i
}
df$id = stringr::str_sub(df$id, 1, 5)

df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=id, y=value, colour=id), size=2) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90))
```

## 9 components

```{r, results='hold', fig.width=4.4, fig.height=8}
hp_temp = hp.mod[[9]]
data.post = cbind(data.sim %>% mutate(clust=cluster_partition(mod$z, hp_temp)), 
                  data.frame(post = apply(mod$z, 1, max)))  %>% tbl_df
print(ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post), size=4, alpha=0.2) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~clust, ncol=3) +
  scale_colour_manual(values=rainbow(K)) )

df = mod$parameters$mean %>% t %>% data.frame %>% setNames(balance_nms) %>%
  mutate(id = 1:K) %>% 
  gather(key='balance', value='value', -id)
for(i in names(hp_temp)){
  df$id[df$id %in% hp_temp[[i]]] = i
}
df$id = stringr::str_sub(df$id, 1, 5)

df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=id, y=value, colour=id), size=2) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90))
```

## 8 components

```{r, results='hold', fig.width=4.4, fig.height=8}
hp_temp = hp.mod[[8]]
data.post = cbind(data.sim %>% mutate(clust=cluster_partition(mod$z, hp_temp)), 
                  data.frame(post = apply(mod$z, 1, max)))  %>% tbl_df
print(ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post), size=4, alpha=0.2) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~clust, ncol=3) +
  scale_colour_manual(values=rainbow(K)) )

df = mod$parameters$mean %>% t %>% data.frame %>% setNames(balance_nms) %>%
  mutate(id = 1:K) %>% 
  gather(key='balance', value='value', -id)
for(i in names(hp_temp)){
  df$id[df$id %in% hp_temp[[i]]] = i
}
df$id = stringr::str_sub(df$id, 1, 5)

df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=id, y=value, colour=id), size=2) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90))
```

## 7 components

```{r, results='hold', fig.width=4.4, fig.height=8}
hp_temp = hp.mod[[7]]
data.post = cbind(data.sim %>% mutate(clust=cluster_partition(mod$z, hp_temp)), 
                  data.frame(post = apply(mod$z, 1, max)))  %>% tbl_df
print(ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post), size=4, alpha=0.2) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~clust, ncol=3) +
  scale_colour_manual(values=rainbow(K)) )

df = mod$parameters$mean %>% t %>% data.frame %>% setNames(balance_nms) %>%
  mutate(id = 1:K) %>% 
  gather(key='balance', value='value', -id)
for(i in names(hp_temp)){
  df$id[df$id %in% hp_temp[[i]]] = i
}
df$id = stringr::str_sub(df$id, 1, 5)

df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=id, y=value, colour=id), size=2) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90))
```

## 6 components

```{r, results='hold', fig.width=4.4, fig.height=8}
hp_temp = hp.mod[[6]]
data.post = cbind(data.sim %>% mutate(clust=cluster_partition(mod$z, hp_temp)), 
                  data.frame(post = apply(mod$z, 1, max)))  %>% tbl_df
print(ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post), size=4, alpha=0.2) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~clust, ncol=3) +
  scale_colour_manual(values=rainbow(K)) )

df = mod$parameters$mean %>% t %>% data.frame %>% setNames(balance_nms) %>%
  mutate(id = 1:K) %>% 
  gather(key='balance', value='value', -id)
for(i in names(hp_temp)){
  df$id[df$id %in% hp_temp[[i]]] = i
}
df$id = stringr::str_sub(df$id, 1, 5)

df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=id, y=value, colour=id), size=2) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90))
```

## 5 components

```{r, results='hold', fig.width=4.4, fig.height=8}
hp_temp = hp.mod[[5]]
data.post = cbind(data.sim %>% mutate(clust=cluster_partition(mod$z, hp_temp)), 
                  data.frame(post = apply(mod$z, 1, max)))  %>% tbl_df
print(ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post), size=4, alpha=0.2) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~clust, ncol=3) +
  scale_colour_manual(values=rainbow(K)) )

df = mod$parameters$mean %>% t %>% data.frame %>% setNames(balance_nms) %>%
  mutate(id = 1:K) %>% 
  gather(key='balance', value='value', -id)
for(i in names(hp_temp)){
  df$id[df$id %in% hp_temp[[i]]] = i
}
df$id = stringr::str_sub(df$id, 1, 5)

df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=id, y=value, colour=id), size=2) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90))
```

## 4 components

```{r, results='hold', fig.width=4.4, fig.height=8}
hp_temp = hp.mod[[4]]
data.post = cbind(data.sim %>% mutate(clust=cluster_partition(mod$z, hp_temp)), 
                  data.frame(post = apply(mod$z, 1, max)))  %>% tbl_df
print(ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post), size=4, alpha=0.2) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~clust, ncol=3) +
  scale_colour_manual(values=rainbow(K)) )

df = mod$parameters$mean %>% t %>% data.frame %>% setNames(balance_nms) %>%
  mutate(id = 1:K) %>% 
  gather(key='balance', value='value', -id)
for(i in names(hp_temp)){
  df$id[df$id %in% hp_temp[[i]]] = i
}
df$id = stringr::str_sub(df$id, 1, 5)

df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=id, y=value, colour=id), size=2) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90))
```

## 3 components

```{r, results='hold', fig.width=4.4, fig.height=8}
hp_temp = hp.mod[[3]]
data.post = cbind(data.sim %>% mutate(clust=cluster_partition(mod$z, hp_temp)), 
                  data.frame(post = apply(mod$z, 1, max)))  %>% tbl_df
print(ggmap(map) + 
  geom_point(data = data.post, 
             aes(x = lon, y = lat, alpha=post), size=4, alpha=0.2) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), 
        axis.text = element_blank(),
        legend.position = 'none') +
  facet_wrap(~clust, ncol=3) +
  scale_colour_manual(values=rainbow(K)) )

df = mod$parameters$mean %>% t %>% data.frame %>% setNames(balance_nms) %>%
  mutate(id = 1:K) %>% 
  gather(key='balance', value='value', -id)
for(i in names(hp_temp)){
  df$id[df$id %in% hp_temp[[i]]] = i
}
df$id = stringr::str_sub(df$id, 1, 5)

df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

ggplot() + geom_point(data=df, aes(x=id, y=value, colour=id), size=2) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90))
```