---
title: "Presentation IAMG 2015"
author: "Marc Comas-Cufí"
date: "27 Jul 2015"
output: 
  html_document: 
    toc: yes
---

# Data presentation

```{r, include=F}
knitr::opts_chunk$set(warning=FALSE, echo=FALSE, comment="   ")
options(width=200)
library(ggmap)
library(ggplot2)
library(tidyr)
library(dplyr)
library(mixpack)
library(gridExtra)
library(Rmixmod)
library(abind)
source('R/plot_functions.R')
load('data/iamg_data.RData')
load('data/iamg_map.RData')
source('functions.R')
```

We worked with a sample of `4,804` observations. 
  
  * The dataset had the following measures:
    * `location`, 
    * `pH`, `conductivity` and 
    * the relative ammount of $(HCO_3, Ca, Cl, Mg, K, Na, SO_4)$.

```{r}
data %>% select(-id, -lon, -lat)
```

  * `27` observation had a zero in component $K$. The main quantiles were:
  
```{r}
data %>% select(HCO3, Ca, Cl, Mg, K, Na, SO4) %>% closure %>% 
  gather(key=component, value=val) %>% 
  group_by(component) %>%
  summarise(
    min = min(val, na.rm=T),
    q1 = quantile(val, 0.25, na.rm=T),
    median = median(val, na.rm=T),
    q3 = quantile(val, 0.75, na.rm=T),
    max = max(val, na.rm=T), 
    zeros = sum(val == 0, na.rm=T))
```
  
  * `27` zeros appeared in part 27,
  * `2` compositions were missing,
  * `539` observation did not have the location well registered

```{r, fig.width=4, fig.height=4}
ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat), alpha=1, col='black') +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank())
```

* We work with a sample of `4,254` observations without zeros, missing data and with registered location.

```{r}
data = data %>% clean_zeros %>% clean_non_located
```

# Classifying the available data

To visualize the chemistry of a water sample, it is common to use the Piper diagram. It is formed with three parts: a ternary diagram in the lower left part showing the cations (amalgamating $Na$ and $K$ in one part), a ternary diagram in the lower right part showing the anions, and a diamond plot in the middle representing the combination of $Ca$ and $Cl$ (when they have been closured with respect cations and anions).

```{r, fig.width=8, fig.height=8}
d.meq <- data %>% compo %>% mgL_to_mmol %>% mmol_to_meq

d.ca = d.meq %>% tern_ca %>% closure
d.an = d.meq %>% tern_an %>% closure 

facie = function(d.ca, d.an) {
  res = rep(as.character(NA), nrow(d.ca))
  res[d.ca$Na.K < 50 & d.an$HCO > 50] = 'Ca-HCO3 Type'
  res[d.ca$Na.K > 50 & d.an$HCO < 50] = 'Na-Cl Type'
  res[d.ca$Na.K < 50 & d.an$HCO < 50 & d.ca$Na.K + d.an$HCO < 50] = 'Ca-Cl Type'
  res[d.ca$Na.K < 50 & d.an$HCO < 50 & d.ca$Na.K + d.an$HCO > 50] = 'Mixed Ca-Mg-Cl Type'
  res[d.ca$Na.K > 50 & d.an$HCO > 50 & d.ca$Na.K + d.an$HCO > 150] = 'Na-HCO3 Type'
  res[d.ca$Na.K > 50 & d.an$HCO > 50 & d.ca$Na.K + d.an$HCO < 150] = 'Mixed Ca-Na-HCO3 Type'
  res
}
piper_data =  with(bind_cols(d.ca, d.an), transform_piper_data(Mg=Mg, Ca=Ca, Cl=Cl, SO4=SO4))
data$facie = facie(d.ca, d.an)
piper_data$facie = facie(d.ca, d.an)
(mgl <- ggplot_piper() + 
  geom_point(data=piper_data %>% mutate(g = facie), 
             aes(x=x,y=y, col=g), size=2))
```

  * To plot the Piper diagram, data sould be transformed to molar equivalent (meq)
  * Once the data is plotted a fixed classification is used to determine the type of water.

```{r, fig.width=9, fig.height=6}
ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat), alpha=0.4, col='black') +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + facet_wrap(~facie)
```

# Compositional approach

We propose to model our data using a finite mixture model. We expres our sample in coordinates using the following balances:

```
b1 = sqrt(4*3/(4+3)) * log(gmean(Ca, Mg, Na, K) / gmean(HCO3, SO4, Cl)),
b2 = sqrt(2*2/(2+2)) * log(gmean(Ca, Mg) / gmean(Na, K)),
b3 = sqrt(1*1/(1+1)) * log(Ca/Mg),
b4 = sqrt(1*1/(1+1)) * log(Na/K),
b5 = sqrt(1*2/(1+2)) * log(gmean(HCO3)/gmean(SO4, Cl)),
b6 = sqrt(1*1/(1+1)) * log(SO4/Cl) 
```

```{r, fig.width=9, fig.height=6}
data.balances = data %>% balance %>% setNames(balance_nms)

ggmap(map) + 
  geom_point(data = cbind(data %>% balance %>% setNames(balance_nms), data %>% select(lon, lat)) %>% 
               gather(key=balance, value=Coordinate, -lon, -lat), 
             aes(x = lon, y = lat, col = Coordinate)) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + facet_wrap(~balance) +
  scale_colour_continuous( low="blue", high="red" ) 
```

```{r}
fit = function(.data, .k, .seed) 
  mixmodCluster(.data, nbCluster = .k, 
                models = mixmodGaussianModel(listModels = 'Gaussian_pk_Lk_Ck'),
                strategy = mixmodStrategy(seed=.seed, nbTry = 5))
```

```{r, eval=F}
mn5 = fit(data.balances, 5, 1)
mn6 = fit(data.balances, 6, 1)
disp_diff = function(mn5, seed){
  sim1 = rmixnorm_solution(4254, mn)

  fit5 = fit(as.data.frame(sim1), 5, seed)
  fit6 = fit(as.data.frame(sim1), 6, seed)
  
  -2*fit5@bestResult@likelihood--2*fit6@bestResult@likelihood
}
test.statistics = sapply(10:29, function(i, mn) disp_diff(mn, i), mn5)
-2*mn5@bestResult@likelihood--2*mn6@bestResult@likelihood
```

## One mixture components - one cluster

```{r, eval=F}
one_more_component = function(K){
  mn.K = fit(data.balances, K, 1)
  mn.K_next = fit(data.balances, K + 1, 1)
  
  disp_diff = function(mn, seed){
    sim1 = rmixnorm_solution(4254, mn)
  
    fit.K = fit(as.data.frame(sim1), mn@nbCluster, seed)
    fit.K_next = fit(as.data.frame(sim1), mn@nbCluster+1, seed)
    
    -2*fit.K@bestResult@likelihood--2*fit.K_next@bestResult@likelihood
  }
  list('obs' = -2*mn.K@bestResult@likelihood--2*mn.K_next@bestResult@likelihood,
       'stat.sample' = sapply(1:25, function(i, mn) disp_diff(mn, i), mn.K))
}

results = lapply(10:15, one_more_component)
sort(c(obs, sample(unlist(test.statistics), 19, replace = F))) == obs
```

Using the test proposed by Aitkin we end up fitting a mixture with $K=14$ components to the dataset. The observations can be classified according to the component more like to be generated by.

```{r}
xlog = function(x) x * log(x)
l_lambda = list(
  'entr' = function(v_tau, a, b) xlog(v_tau[a] + v_tau[b]) - xlog(v_tau[a]) - xlog(v_tau[b]),
  'demp' = function(v_tau, a, b) if(which.max(v_tau) == b) 1 else 0,
  'demp.mod' = function(v_tau, a, b) v_tau[b] * (v_tau[a] + v_tau[b])^-1,
  'coda' = function(v_tau, a, b) log(v_tau[b] / v_tau[a]),
  'coda.norm' = function(v_tau, a, b) -log(v_tau[b] / v_tau[a])^2,
  'prop' = function(v_tau, a, b) v_tau[b] )

# Weitghing functions
l_omega = list(
  'cnst' = function(v_tau, a) 1,
  'prop' = function(v_tau, a) v_tau[a],
  'dich' = function(v_tau, a) if(which.max(v_tau) == a) 1 else 0
)
```

```{r}
K = 14
mod = fit(data.balances, K, 1)
POST = mod@bestResult@proba
POST[POST==0] = .Machine$double.xmin
hp = get_hierarchical_partition(POST,
                                omega = l_omega[['prop']],
                                lambda = l_lambda[['coda.norm']])
hp.mod = hp
df = mod@bestResult@parameters@mean %>% data.frame %>% tbl_df %>% 
  setNames(balance_nms) %>%
  mutate(g0 = sprintf('%02d', 1:mod@bestResult@nbCluster)) %>% 
  gather(key='balance', value='value', -g0)
df.mean = df %>% group_by(balance) %>% summarise(value = mean(value))

data$g0 = sprintf('%02d', mod@bestResult@partition) 
```

```{r, results='hold', fig.width=4.4, fig.height=8}
HP  = hp.mod[[14]]
hp_temp = lapply(HP, function(i) sprintf('%02d', i))

df$g = ''
for(i in names(hp_temp)){
  df$g[df$g0 %in% hp_temp[[i]]] = i
}

lvls = df %>% filter(balance == '(Ca·Mg·Na·K)/(HCO3·SO4·Cl)') %>% 
                  group_by(g) %>% 
                  summarise( value.m = mean(value) ) %>% 
                  arrange(value.m) %>% select(g) %>% .[[1]]
data$g = factor(cluster_partition(mod@bestResult@proba, HP), levels = lvls)
df$g = factor(df$g, lvls)

ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat, col = g0), alpha=0.4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + 
  facet_wrap(~g, ncol=3) + theme(legend.position = 'none')

ggplot() + 
  geom_point(data=df, 
             aes(x=g, y=value, colour=g0), size=3) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90)) + 
  scale_x_discrete(breaks=lvls, labels=ifelse(nchar(lvls) > 5, paste0(substr(lvls, 1, 4), '...'), lvls))
```

## Different mixture components - one cluster

After fitting a mixture with 14 components, we continue by merging sequantially those components more similar. We have used functions to define the similarity score:

```
omega = function(v_tau, a, b) v_tau[b]
lambda = function(v_tau, a, b) -log(v_tau[b] / v_tau[a])^2
```

### With K = 13 components

```{r, results='hold', fig.width=4.4, fig.height=8}
HP  = hp.mod[[13]]
hp_temp = lapply(HP, function(i) sprintf('%02d', i))

df$g = ''
for(i in names(hp_temp)){
  df$g[df$g0 %in% hp_temp[[i]]] = i
}

lvls = df %>% filter(balance == '(Ca·Mg·Na·K)/(HCO3·SO4·Cl)') %>% 
                  group_by(g) %>% 
                  summarise( value.m = mean(value) ) %>% 
                  arrange(value.m) %>% select(g) %>% .[[1]]
data$g = factor(cluster_partition(mod@bestResult@proba, HP), levels = lvls)
df$g = factor(df$g, lvls)

ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat, col = g0), alpha=0.4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + 
  facet_wrap(~g, ncol=3) + theme(legend.position = 'none')

ggplot() + 
  geom_point(data=df, 
             aes(x=g, y=value, colour=g0), size=3) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90)) + 
  scale_x_discrete(breaks=lvls, labels=ifelse(nchar(lvls) > 5, paste0(substr(lvls, 1, 4), '...'), lvls))
```

### With K = 12 components

```{r, results='hold', fig.width=4.4, fig.height=8}
HP  = hp.mod[[12]]
hp_temp = lapply(HP, function(i) sprintf('%02d', i))

df$g = ''
for(i in names(hp_temp)){
  df$g[df$g0 %in% hp_temp[[i]]] = i
}

lvls = df %>% filter(balance == '(Ca·Mg·Na·K)/(HCO3·SO4·Cl)') %>% 
                  group_by(g) %>% 
                  summarise( value.m = mean(value) ) %>% 
                  arrange(value.m) %>% select(g) %>% .[[1]]
data$g = factor(cluster_partition(mod@bestResult@proba, HP), levels = lvls)
df$g = factor(df$g, lvls)

ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat, col = g0), alpha=0.4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + 
  facet_wrap(~g, ncol=3) + theme(legend.position = 'none')

ggplot() + 
  geom_point(data=df, 
             aes(x=g, y=value, colour=g0), size=3) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90)) + 
  scale_x_discrete(breaks=lvls, labels=ifelse(nchar(lvls) > 5, paste0(substr(lvls, 1, 4), '...'), lvls))
```

### With K = 11 components

```{r, results='hold', fig.width=4.4, fig.height=8}
HP  = hp.mod[[11]]
hp_temp = lapply(HP, function(i) sprintf('%02d', i))

df$g = ''
for(i in names(hp_temp)){
  df$g[df$g0 %in% hp_temp[[i]]] = i
}

lvls = df %>% filter(balance == '(Ca·Mg·Na·K)/(HCO3·SO4·Cl)') %>% 
                  group_by(g) %>% 
                  summarise( value.m = mean(value) ) %>% 
                  arrange(value.m) %>% select(g) %>% .[[1]]
data$g = factor(cluster_partition(mod@bestResult@proba, HP), levels = lvls)
df$g = factor(df$g, lvls)

ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat, col = g0), alpha=0.4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + 
  facet_wrap(~g, ncol=3) + theme(legend.position = 'none')

ggplot() + 
  geom_point(data=df, 
             aes(x=g, y=value, colour=g0), size=3) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90)) + 
  scale_x_discrete(breaks=lvls, labels=ifelse(nchar(lvls) > 5, paste0(substr(lvls, 1, 4), '...'), lvls))
```


### With K = 10 components

```{r, results='hold', fig.width=4.4, fig.height=8}
HP  = hp.mod[[10]]
hp_temp = lapply(HP, function(i) sprintf('%02d', i))

df$g = ''
for(i in names(hp_temp)){
  df$g[df$g0 %in% hp_temp[[i]]] = i
}

lvls = df %>% filter(balance == '(Ca·Mg·Na·K)/(HCO3·SO4·Cl)') %>% 
                  group_by(g) %>% 
                  summarise( value.m = mean(value) ) %>% 
                  arrange(value.m) %>% select(g) %>% .[[1]]
data$g = factor(cluster_partition(mod@bestResult@proba, HP), levels = lvls)
df$g = factor(df$g, lvls)

ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat, col = g0), alpha=0.4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + 
  facet_wrap(~g, ncol=3) + theme(legend.position = 'none')

ggplot() + 
  geom_point(data=df, 
             aes(x=g, y=value, colour=g0), size=3) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90)) + 
  scale_x_discrete(breaks=lvls, labels=ifelse(nchar(lvls) > 5, paste0(substr(lvls, 1, 4), '...'), lvls))
```


### With K = 9 components

```{r, results='hold', fig.width=4.4, fig.height=8}
HP  = hp.mod[[9]]
hp_temp = lapply(HP, function(i) sprintf('%02d', i))

df$g = ''
for(i in names(hp_temp)){
  df$g[df$g0 %in% hp_temp[[i]]] = i
}

lvls = df %>% filter(balance == '(Ca·Mg·Na·K)/(HCO3·SO4·Cl)') %>% 
                  group_by(g) %>% 
                  summarise( value.m = mean(value) ) %>% 
                  arrange(value.m) %>% select(g) %>% .[[1]]
data$g = factor(cluster_partition(mod@bestResult@proba, HP), levels = lvls)
df$g = factor(df$g, lvls)

ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat, col = g0), alpha=0.4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + 
  facet_wrap(~g, ncol=3) + theme(legend.position = 'none')

ggplot() + 
  geom_point(data=df, 
             aes(x=g, y=value, colour=g0), size=3) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90)) + 
  scale_x_discrete(breaks=lvls, labels=ifelse(nchar(lvls) > 5, paste0(substr(lvls, 1, 4), '...'), lvls))
```

### With K = 8 components

```{r, results='hold', fig.width=4.4, fig.height=8}
HP  = hp.mod[[8]]
hp_temp = lapply(HP, function(i) sprintf('%02d', i))

df$g = ''
for(i in names(hp_temp)){
  df$g[df$g0 %in% hp_temp[[i]]] = i
}

lvls = df %>% filter(balance == '(Ca·Mg·Na·K)/(HCO3·SO4·Cl)') %>% 
                  group_by(g) %>% 
                  summarise( value.m = mean(value) ) %>% 
                  arrange(value.m) %>% select(g) %>% .[[1]]
data$g = factor(cluster_partition(mod@bestResult@proba, HP), levels = lvls)
df$g = factor(df$g, lvls)

ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat, col = g0), alpha=0.4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + 
  facet_wrap(~g, ncol=3) + theme(legend.position = 'none')

ggplot() + 
  geom_point(data=df, 
             aes(x=g, y=value, colour=g0), size=3) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90)) + 
  scale_x_discrete(breaks=lvls, labels=ifelse(nchar(lvls) > 5, paste0(substr(lvls, 1, 4), '...'), lvls))
```

### With K = 7 components

```{r, results='hold', fig.width=4.4, fig.height=8}
HP  = hp.mod[[7]]
hp_temp = lapply(HP, function(i) sprintf('%02d', i))

df$g = ''
for(i in names(hp_temp)){
  df$g[df$g0 %in% hp_temp[[i]]] = i
}

lvls = df %>% filter(balance == '(Ca·Mg·Na·K)/(HCO3·SO4·Cl)') %>% 
                  group_by(g) %>% 
                  summarise( value.m = mean(value) ) %>% 
                  arrange(value.m) %>% select(g) %>% .[[1]]
data$g = factor(cluster_partition(mod@bestResult@proba, HP), levels = lvls)
df$g = factor(df$g, lvls)

ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat, col = g0), alpha=0.4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + 
  facet_wrap(~g, ncol=3) + theme(legend.position = 'none')

ggplot() + 
  geom_point(data=df, 
             aes(x=g, y=value, colour=g0), size=3) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90)) + 
  scale_x_discrete(breaks=lvls, labels=ifelse(nchar(lvls) > 5, paste0(substr(lvls, 1, 4), '...'), lvls))
```

### With K = 6 components

```{r, results='hold', fig.width=4.4, fig.height=8}
HP  = hp.mod[[6]]
hp_temp = lapply(HP, function(i) sprintf('%02d', i))

df$g = ''
for(i in names(hp_temp)){
  df$g[df$g0 %in% hp_temp[[i]]] = i
}

lvls = df %>% filter(balance == '(Ca·Mg·Na·K)/(HCO3·SO4·Cl)') %>% 
                  group_by(g) %>% 
                  summarise( value.m = mean(value) ) %>% 
                  arrange(value.m) %>% select(g) %>% .[[1]]
data$g = factor(cluster_partition(mod@bestResult@proba, HP), levels = lvls)
df$g = factor(df$g, lvls)

ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat, col = g0), alpha=0.4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + 
  facet_wrap(~g, ncol=3) + theme(legend.position = 'none')

ggplot() + 
  geom_point(data=df, 
             aes(x=g, y=value, colour=g0), size=3) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90)) + 
  scale_x_discrete(breaks=lvls, labels=ifelse(nchar(lvls) > 5, paste0(substr(lvls, 1, 4), '...'), lvls))
```

### With K = 5 components

```{r, results='hold', fig.width=4.4, fig.height=8}
HP  = hp.mod[[5]]
hp_temp = lapply(HP, function(i) sprintf('%02d', i))

df$g = ''
for(i in names(hp_temp)){
  df$g[df$g0 %in% hp_temp[[i]]] = i
}

lvls = df %>% filter(balance == '(Ca·Mg·Na·K)/(HCO3·SO4·Cl)') %>% 
                  group_by(g) %>% 
                  summarise( value.m = mean(value) ) %>% 
                  arrange(value.m) %>% select(g) %>% .[[1]]
data$g = factor(cluster_partition(mod@bestResult@proba, HP), levels = lvls)
df$g = factor(df$g, lvls)

ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat, col = g0), alpha=0.4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + 
  facet_wrap(~g, ncol=3) + theme(legend.position = 'none')

ggplot() + 
  geom_point(data=df, 
             aes(x=g, y=value, colour=g0), size=3) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90)) + 
  scale_x_discrete(breaks=lvls, labels=ifelse(nchar(lvls) > 5, paste0(substr(lvls, 1, 4), '...'), lvls))
```

### With K = 4 components

```{r, results='hold', fig.width=4.4, fig.height=8}
HP  = hp.mod[[4]]
hp_temp = lapply(HP, function(i) sprintf('%02d', i))

df$g = ''
for(i in names(hp_temp)){
  df$g[df$g0 %in% hp_temp[[i]]] = i
}

lvls = df %>% filter(balance == '(Ca·Mg·Na·K)/(HCO3·SO4·Cl)') %>% 
                  group_by(g) %>% 
                  summarise( value.m = mean(value) ) %>% 
                  arrange(value.m) %>% select(g) %>% .[[1]]
data$g = factor(cluster_partition(mod@bestResult@proba, HP), levels = lvls)
df$g = factor(df$g, lvls)

ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat, col = g0), alpha=0.4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + 
  facet_wrap(~g, ncol=3) + theme(legend.position = 'none')

ggplot() + 
  geom_point(data=df, 
             aes(x=g, y=value, colour=g0), size=3) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90)) + 
  scale_x_discrete(breaks=lvls, labels=ifelse(nchar(lvls) > 5, paste0(substr(lvls, 1, 4), '...'), lvls))
```

### With K = 3 components

```{r, results='hold', fig.width=4.4, fig.height=8}
HP  = hp.mod[[3]]
hp_temp = lapply(HP, function(i) sprintf('%02d', i))

df$g = ''
for(i in names(hp_temp)){
  df$g[df$g0 %in% hp_temp[[i]]] = i
}

lvls = df %>% filter(balance == '(Ca·Mg·Na·K)/(HCO3·SO4·Cl)') %>% 
                  group_by(g) %>% 
                  summarise( value.m = mean(value) ) %>% 
                  arrange(value.m) %>% select(g) %>% .[[1]]
data$g = factor(cluster_partition(mod@bestResult@proba, HP), levels = lvls)
df$g = factor(df$g, lvls)

ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat, col = g0), alpha=0.4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + 
  facet_wrap(~g, ncol=3) + theme(legend.position = 'none')

ggplot() + 
  geom_point(data=df, 
             aes(x=g, y=value, colour=g0), size=3) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90)) + 
  scale_x_discrete(breaks=lvls, labels=ifelse(nchar(lvls) > 5, paste0(substr(lvls, 1, 4), '...'), lvls))
```

### With K = 2 components

```{r, results='hold', fig.width=4.4, fig.height=8}
HP  = hp.mod[[2]]
hp_temp = lapply(HP, function(i) sprintf('%02d', i))

df$g = ''
for(i in names(hp_temp)){
  df$g[df$g0 %in% hp_temp[[i]]] = i
}

lvls = df %>% filter(balance == '(Ca·Mg·Na·K)/(HCO3·SO4·Cl)') %>% 
                  group_by(g) %>% 
                  summarise( value.m = mean(value) ) %>% 
                  arrange(value.m) %>% select(g) %>% .[[1]]
data$g = factor(cluster_partition(mod@bestResult@proba, HP), levels = lvls)
df$g = factor(df$g, lvls)

ggmap(map) + 
  geom_point(data = data, aes(x = lon, y = lat, col = g0), alpha=0.4) +
  xlab(NULL) + ylab(NULL) +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) + 
  facet_wrap(~g, ncol=3) + theme(legend.position = 'none')

ggplot() + 
  geom_point(data=df, 
             aes(x=g, y=value, colour=g0), size=3) +
  geom_hline(data=df.mean, aes(yintercept=value), col='blue') + 
  facet_wrap(~balance, ncol=1, scale='free') + 
  theme_bw() + theme(legend.position = 'none', 
                     axis.text.x=element_text(angle=-90)) + 
  scale_x_discrete(breaks=lvls, labels=ifelse(nchar(lvls) > 5, paste0(substr(lvls, 1, 4), '...'), lvls))
```

