```{r, include=FALSE}
library(ggplot2)
library(ggmap)
library(grid)
library(dplyr)
library(reshape2)
library(ggtern)
library(stringr)
library(mixpack)
library(flexmix)
if(!exists('ROOT')) ROOT = getwd()
load(sprintf('%s/data/clean_data.RData', ROOT))
```

Descriptives
============

* [(Ca,Mg,Na)-composition](#comp1)
  * [Original composition](#comp1_tern)
  * [ILR representation](#comp1_ilr)
  * [Plotting the Facies into a map](#comp1_facie)
  * [Plotting the CLR transform coordinates into a map](#comp1_clr)
* [(SO4,HCO3,Cl)-composition](#comp2)
  * [Original composition](#comp2_tern)
  * [ILR representation](#comp2_ilr)
  * [Plotting the Facies into a map](#comp2_facie)
  * [Plotting the CLR transform coordinates into a map](#comp2_clr)

## <a name="comp1" />(Ca,Mg,Na)-composition

```{r}
df1.coda = data %>% 
  mutate(
    'Facie' = sapply(str_split(data$Facies, '-'), function(v) v[1])
    ) %>%
  select(x, y, Facie, Ca, Mg, Na)
df1.clr = df1.coda %>% 
  mutate(
    'clrCa' = log(Ca/(Ca*Mg*Na)^(1/3)),
    'clrMg' = log(Mg/(Ca*Mg*Na)^(1/3)),
    'clrNa' = log(Na/(Ca*Mg*Na)^(1/3))
    ) %>% 
  select(x, y, Facie, clrCa, clrMg, clrNa)
df1.ilr = ilr_coordinates(df1.coda %>% select(Ca, Mg, Na) )  %>% cbind(df1.coda %>% select(Facie))
## Variation array
df1.coda %>% select(Ca, Mg, Na) %>% 
  melt(id.vars = c('Ca', 'Mg', 'Na'), measure.vars = c('Ca', 'Mg', 'Na'), 
       value.name = 'val.row', variable.name = 'var.row') %>%
  melt(id.vars = c('Ca', 'Mg', 'Na', 'var.row', 'val.row'), measure.vars = c('Ca', 'Mg', 'Na'), 
       value.name = 'val.col', variable.name = 'var.col') %>% 
  group_by(var.row, var.col) %>% summarize(
  'm' = mean(log(val.row/val.col)),
  'v' = var(log(val.row/val.col)),
  'l' = sprintf("  %5.2f (%5.2f)", m, v)) %>% dcast(var.row~var.col, value.var = 'l')
```

### <a name="comp1_tern" /> Original composition

```{r}
ggtern(data = df1.coda, aes(x=Ca, Mg, Na)) + geom_point(aes(col=Facie), alpha=0.25) + geom_density2d()
```

### <a name="comp1_ilr" /> ILR representation

```{r, fig.width=7, fig.height=7}
ggplot( data = df1.ilr ) +
  geom_point(aes(x=coord.1, y=coord.2, col=Facie), alpha=0.5) + geom_density2d(aes(x=coord.1, y=coord.2))
```

### <a name="comp1_facie" /> Plotting the Facies into a map

```{r, fig.width=14, fig.height=5, warning=FALSE}
df1 = df1.clr %>% melt(id.vars = c('x', 'y', 'Facie'))
ggmap(map) + 
  geom_point(data = df1, aes(x = x, y = y, col=Facie), size=3,  alpha = .5) +
  xlab(NULL) + ylab(NULL) + facet_wrap(~variable, nrow=1) +
  theme(axis.ticks = element_blank(), axis.text = element_blank())
```

### <a name="comp1_clr" /> Plotting the CLR transform coordinates into a map

```{r, fig.width=14, fig.height=5, warning=FALSE}
ggmap(map) + 
  geom_point(data = df1, aes(x = x, y = y, color=value), size=3,  alpha = .5) +
  xlab(NULL) + ylab(NULL) + facet_wrap(~variable, nrow=1) +
  theme(axis.ticks = element_blank(), axis.text = element_blank())
```

## <a name="comp2" />(SO4,HCO3,Cl)-composition

```{r}
df2.coda <- data %>%
  mutate(
    'Facie' = sapply(str_split(data$Facies, '-'), function(v) v[2])
    ) %>% 
  select(x, y, Facie, SO4, HCO3, Cl)
df2.clr = df2.coda %>% 
  mutate(
    'clrSO4' = log(SO4/(SO4*HCO3*Cl)^(1/3)),
    'clrHCO3' = log(HCO3/(SO4*HCO3*Cl)^(1/3)),
    'clrCl' = log(Cl/(SO4*HCO3*Cl)^(1/3))
    ) %>% 
  select(x, y, Facie, clrSO4, clrHCO3, clrCl)
df2.ilr = ilr_coordinates(df2.coda %>% select(SO4, HCO3, Cl))  %>% cbind(df2.coda %>% select(Facie))
## Variation array
df2.coda %>% select(SO4, HCO3, Cl) %>% 
  melt(id.vars = c('SO4', 'HCO3', 'Cl'), measure.vars = c('SO4', 'HCO3', 'Cl'), 
       value.name = 'val.row', variable.name = 'var.row') %>%
  melt(id.vars = c('SO4', 'HCO3', 'Cl', 'var.row', 'val.row'), measure.vars = c('SO4', 'HCO3', 'Cl'), 
       value.name = 'val.col', variable.name = 'var.col') %>% 
  group_by(var.row, var.col) %>% summarize(
  'm' = mean(log(val.row/val.col)),
  'v' = var(log(val.row/val.col)),
  'l' = sprintf("  %5.2f (%5.2f)", m, v)) %>% dcast(var.row~var.col, value.var = 'l')
```

### <a name="comp2_tern" /> Original composition

```{r}
ggtern(data = df2.coda, aes(x=SO4, HCO3, Cl)) + geom_point(aes(col=Facie), alpha=0.25) + geom_density2d()
```

### <a name="comp2_ilr" /> ILR representation

```{r, fig.width=7, fig.height=7}
ggplot( data = df2.ilr ) +
  geom_point(aes(x=coord.1, y=coord.2, col=Facie), alpha=0.5) + geom_density2d(aes(x=coord.1, y=coord.2))
```

### <a name="comp2_facie" /> Plotting the Facies into a map

```{r, fig.width=14, fig.height=5, warning=FALSE}
df2 = df2.clr %>% melt(id.vars = c('x', 'y', 'Facie'))
ggmap(map) + 
  geom_point(data = df2, aes(x = x, y = y, col=Facie), size=3,  alpha = .5) +
  xlab(NULL) + ylab(NULL) + facet_wrap(~variable, nrow=1) +
  theme(axis.ticks = element_blank(), axis.text = element_blank())
```

### <a name="comp2_clr" /> Plotting the CLR transform coordinates into a map

```{r, fig.width=14, fig.height=5, warning=FALSE}
ggmap(map) + 
  geom_point(data = df2, aes(x = x, y = y, color=value), size=3,  alpha = .5) +
  xlab(NULL) + ylab(NULL) + facet_wrap(~variable, nrow=1) +
  theme(axis.ticks = element_blank(), axis.text = element_blank())
```


## Different approaches

### Mixture regression

```{r}
set.seed(1024)
```

#### (Ca,Mg,Na)-composition

```{r}
summary( m1 <- flexmix(coord.1~coord.2, data = df1.ilr, k = 10) )
df1.reg = df1.coda %>% cbind('cluster' = as.factor(m1@cluster))
```

```{r}
ggtern(data = df1.reg, aes(x=Ca, Mg, Na)) + geom_point(aes(col=cluster))
```

```{r, fig.width=14, fig.height=5, warning=FALSE}
df1 = df1.reg %>% melt(id.vars = c('x', 'y', 'Facie', 'cluster'))
ggmap(map) + 
  geom_point(data = df1, aes(x = x, y = y, col=cluster), size=3,  alpha = .5) +
  xlab(NULL) + ylab(NULL) + facet_wrap(~variable, nrow=1) +
  theme(axis.ticks = element_blank(), axis.text = element_blank())
```

#### (SO4,HCO3,Cl)-composition

```{r}
summary( m2 <- flexmix(coord.1~coord.2, data = df2.ilr, k = 10) )
df2.reg = df2.coda %>% cbind('cluster' = as.factor(m2@cluster))
```

```{r}
ggtern(data = df2.reg, aes(x=SO4, HCO3, Cl)) + geom_point(aes(col=cluster))
```

```{r, fig.width=14, fig.height=5, warning=FALSE}
df2 = df2.reg %>% melt(id.vars = c('x', 'y', 'Facie', 'cluster'))
ggmap(map) + 
  geom_point(data = df2, aes(x = x, y = y, col=cluster), size=3,  alpha = .5) +
  xlab(NULL) + ylab(NULL) + facet_wrap(~variable, nrow=1) +
  theme(axis.ticks = element_blank(), axis.text = element_blank())
```