```{r, include=FALSE}
library(ggplot2)
library(ggmap)
library(grid)
library(dplyr)
library(reshape2)
library(ggtern)
library(stringr)
library(mixpack)
library(flexmix)
library(gridExtra)
if(!exists('ROOT')) ROOT = getwd()
load(sprintf('%s/data/clean_data.RData', ROOT))
source(sprintf('%s/R/plot_functions.R', ROOT))

df = data %>% mutate(
  HCO3 = Bicarbonati_mg.L_, 
  Ca = Calcio_mg.L_, 
  Cl = Cloruri_mg.L_,
  Mg = Magnesio_mg.L_, 
  K = Potassio_mg.L_, 
  Na = Sodio_mg.L_,
  SO4 = Solfati_mg.L_) %>% select(HCO3, Ca, Cl, Mg, K, Na, SO4) %>% subset() %>% tbl_df
```

The piper plot is not coherent under perturbations
==================================================

Here, we compare the piper plots using Mg/L, mmoL, mEq. We see the representation changes with respect to the units used.

* [Piper plot using Mg/L](#mgl)
* [Piper plot using mmol](#mmol)
* [Piper plot using meq](#meq)
* [Comparing the piper plots](#comparison)

### <a name="mgl" />Piper plot using Mg/L

#### Sample in Mg/L

```{r, echo=FALSE}
df
```

#### Cations and anions

* Cations (closured)

```{r, echo=FALSE}
( df.cat <- df %>% mutate(
  total = Ca + Mg + Na + K,
  Ca = Ca / total * 100,
  Mg = Mg / total * 100,
  Na_K = (Na + K) / total * 100) %>% select(Ca, Mg, Na_K) ) %>% round(2)
```

* Anions (closured)

```{r, echo=FALSE}
( df.ani <- df %>% mutate(
  total = Cl + SO4 + HCO3,
  Cl = Cl / total * 100,
  SO4 = SO4 / total * 100,
  HCO3 = HCO3 / total * 100) %>% select(Cl, SO4, HCO3) %>% round(2) )
```

* Piper plot

```{r, echo=FALSE, fig.width=9, fig.height=9}
facie = function(.data) names(.data)[apply(.data, 1, which.max)]
piper_data <- transform_piper_data(Ca=df.cat$Ca, Mg = df.cat$Mg, Cl=df.ani$Cl, SO4 = df.ani$SO4)
piper_data$facie = sprintf("%s-%s", facie(df.cat), facie(df.ani))
(mgl <- ggplot_piper() + geom_point(data=piper_data, aes(x=x,y=y, col=facie), size=2))
```

### <a name="mmol" />Piper plot using mmol

Dividing factor from Mg/L to mmol

```{r, echo=FALSE}
( conv <- c(HCO3=61.02, Ca=40.078, Cl=35.4527, Mg=24.305, K=39.0983, Na=22.989, SO4=96.06) )
```

#### Sample in mmol

```{r, echo=FALSE}
for(v in names(df)){
  df[,v] = df[,v] / conv[v]
}
df %>% round(2)
```

#### Cations and anions

* Cations (closured)

```{r, echo=FALSE}
( df.cat <- df %>% mutate(
  total = Ca + Mg + Na + K,
  Ca = Ca / total * 100,
  Mg = Mg / total * 100,
  Na_K = (Na + K) / total * 100) %>% select(Ca, Mg, Na_K) ) %>% round(2)
```

* Anions (closured)

```{r, echo=FALSE}
( df.ani <- df %>% mutate(
  total = Cl + SO4 + HCO3,
  Cl = Cl / total * 100,
  SO4 = SO4 / total * 100,
  HCO3 = HCO3 / total * 100) %>% select(Cl, SO4, HCO3) %>% round(2) )
```

* Piper plot

```{r, echo=FALSE, fig.width=9, fig.height=9}
facie = function(.data) names(.data)[apply(.data, 1, which.max)]
piper_data <- transform_piper_data(Ca=df.cat$Ca, Mg = df.cat$Mg, Cl=df.ani$Cl, SO4 = df.ani$SO4)
piper_data$facie = sprintf("%s-%s", facie(df.cat), facie(df.ani))
(mmol <- ggplot_piper() + geom_point(data=piper_data, aes(x=x,y=y, col=facie), size=2))
```

### <a name="meq" />Piper plot using meq

Multiplying factor from mmol to meq

```{r, echo=FALSE}
( conv <- c(HCO3=1, Ca=2, Cl=1, Mg=2, K=1, Na=1, SO4=2) )
```

#### Sample in mmol

```{r, echo=FALSE}
for(v in names(df)){
  df[,v] = df[,v] * conv[v]
}
df %>% round(2)
```

#### Cations and anions

* Cations (closured)

```{r, echo=FALSE}
( df.cat <- df %>% mutate(
  total = Ca + Mg + Na + K,
  Ca = Ca / total * 100,
  Mg = Mg / total * 100,
  Na_K = (Na + K) / total * 100) %>% select(Ca, Mg, Na_K) ) %>% round(2)
```

* Anions (closured)

```{r, echo=FALSE}
( df.ani <- df %>% mutate(
  total = Cl + SO4 + HCO3,
  Cl = Cl / total * 100,
  SO4 = SO4 / total * 100,
  HCO3 = HCO3 / total * 100) %>% select(Cl, SO4, HCO3) %>% round(2) )
```

* Piper plot

```{r, echo=FALSE, fig.width=9, fig.height=9}
facie = function(.data) names(.data)[apply(.data, 1, which.max)]
piper_data <- transform_piper_data(Ca=df.cat$Ca, Mg = df.cat$Mg, Cl=df.ani$Cl, SO4 = df.ani$SO4)
piper_data$facie = sprintf("%s-%s", facie(df.cat), facie(df.ani))
(meq <- ggplot_piper() + geom_point(data=piper_data, aes(x=x,y=y, col=facie), size=2))
```

### <a name="comparison" />Comparing the piper plots

* Mg/L

```{r, echo=FALSE, fig.width=9, fig.height=9}
mgl
```

* mmol

```{r, echo=FALSE, fig.width=9, fig.height=9}
mmol
```

* meq

```{r, echo=FALSE, fig.width=9, fig.height=9}
meq
```